### 2015-06-21 ms.tex
added final paragraph of intro listing sections of the paper

\documentclass{emulateapj}

\usepackage[dvipdf]{epsfig} 
\usepackage[dvips]{rotating}
\usepackage{subfigure}
\usepackage{mathrsfs}
\usepackage{amsmath}

\newcommand{\IRAS}{{\it IRAS}}
\newcommand{\HERSCHEL}{{\it Herschel}}
\newcommand{\SPITZER}{{\it Spitzer}}
\newcommand{\PLANCK}{{\it Planck}}
\newcommand{\AKARI}{{\it AKARI}}
\newcommand{\COBE}{{\it COBE}}
\newcommand{\WISE}{{\it WISE}}

\bibpunct{(}{)}{;}{a}{}{,}

\shorttitle{\PLANCK~2015 Dust Modeling}

\shortauthors{Meisner \& Finkbeiner}

\begin{document}

\title{Modeling thermal dust emission in the {\it PLANCK} 2015 release maps}
\author{Aaron M. Meisner\altaffilmark{1,2}}
\author{Douglas P. Finkbeiner\altaffilmark{1,2}}
\altaffiltext{1}{Department of Physics, Harvard University, 17 Oxford Street, 
Cambridge, MA 02138, USA; ameisner@fas.harvard.edu}
\altaffiltext{2}{Harvard-Smithsonian Center for Astrophysics, 60 Garden St, 
Cambridge, MA 02138, USA; dfinkbeiner@cfa.harvard.edu}

\begin{abstract}

Based on the \PLANCK~2015 data release maps \citep{planck2015}, we present a 
thermal dust emission analysis analogous to that of \cite{meisner15}. We 
isolate thermal dust emission in the \PLANCK~2015 frequency maps, and present 
full-sky models of the unpolarized thermal dust emission using several 
paremeterizations of the dust spectrum, including single modified blackbody and
two-component models. The qualitative conclusions of \cite{meisner15} remain 
unchanged. In particular, the zodiacal light still represents the dominant 
systematic problem of our dust model on large angular scales, and the cosmic 
infrared background anisotropy remains the dominant systematic problem on small
angular scales.

% mention frequency range 100-3000 GHz ?
% cite the FDS99 model in the intro as well ?
% mention that this isn't just a Planck-based SED, but includes DIRBE/IRAS ?
% mention that this is first ever thermal dust model based on Planck 2015 ?

\end{abstract}

\keywords{infrared: ISM, submillimeter: ISM, dust, extinction}

\section{Introduction}
% brief importance of dust
Interstellar dust in our Galaxy represents a crucial and pervasive foreground
for astronomical observations over an extremely wide range of photon energies.
In the optical, ultraviolet and near-infrared, extinction by intervening dust
attenuates and reddens the observed signal from background stars, galaxies and
transients. At millimeter wavelengths, thermal dust emission contaminates
observations of the cosmic microwave background (CMB).

% iras first revolution in mapping ISM cirrus
While the dust in general lies within a disk a few hundred parsecs in scale 
height centered about the Galactic midplane, its detailed distribution is very 
patchy and cloudlike, as first revealed in far-infrared thermal emission by the
groundbreaking IRAS mission. Thus, it is of great importance to map 
interstellar dust over the entire sky, with high angular resolution and 
sensitivity.

% planck more recent breakthrough in mapping dust emission due to high
%     angular resolution / sensitivity of broad range of frequencies
The Planck mission (cite 2013 and 2015 release overview papers) has recently 
provided a breakthrough in full-sky mapping of emission from interstellar dust,
by obtaining relatively artifact free full-sky maps at XX frequencies from QQ 
to PP GHz, with a superb combination of near-IRAS resolution from QQ-PP and 
high sensitivity. Although these \PLANCK~maps contain various emission 
components, including the CMB, zodiacal light,and molecular emission, it is 
nevertheless possible to isolate exquisite maps of thermal emission from 
interstellar dust grains.
% mention the existing dust data products, say they're all very useful for
%     many purposes
Several full-sky data products currently exist for predicting emission and/or
extinction by interstellar dust. SFD - mention DIRBE temperature limitation.
Planck allows us to get around temperature limitation. One sentence overview of
Planck team's model. One or two sentences about Meisner+Finkbeiner modeling.

% but say that they all still are not perfect, and we want to investigate
%     what effect swapping in the latest planck release has on the dust model
% and we also just want to have a model with data from latest Planck release,
%     which has never been done by anyone before
But none of these are yet perfect. Mention CIBA and zodi as examples from
Meisner + Finkbeiner. Also, no Planck 2015 release thermal dust model yet
exists at all.

% list of the sections
Much of our detailed methodology is identical to that of Meisner/Finkbeiner 
2015. As a result, throughout this paper, we mainly emphasize any notable 
differences relative, and otherwise refer the reader to Meisner/Finkbeiner 
2015 for thorough details.

In $\S$\ref{sec:data} we introduce the data used throughout this study. In
$\S$\ref{sec:prepro} we describe our preprocessing of the \PLANCK~maps to
isolate thermal emission from Galactic dust. In $\S$\ref{sec:modeling} we
explain the two-component emission model we apply to the \PLANCK-based dust
SED. In $\S$\ref{sec:bpcorr}, we discuss the details of predicting
\PLANCK~observations based on this dust model. In $\S$\ref{sec:global} we
derive constraints on our model's global parameters in light of the \PLANCK~HFI
maps. In $\S$\ref{sec:fitting} we detail the Markov chain Monte Carlo (MCMC)
method with which we have estimated the spatially varying parameters of our
model. In $\S$\ref{sec:ebv} we calibrate our derived optical depth to reddening
at optical wavelengths. In $\S$\ref{sec:em_compare} we compare our
two-component thermal dust emission predictions to those of \cite{planckdust}.
In $\S$\ref{sec:release} we present the full-sky maps of dust temperature and
optical depth we have obtained, and conclude in $\S$\ref{sec:conclusion}.

\section{Data}
\label{sec:data}

\section{Preprocessing}
\label{sec:prepro}

\subsection{CMB Anisotropy Removal}
\label{sec:cmb}

\subsection{Compact Sources}
\label{sec:ptsrc}

\subsection{Smoothing}
\label{sec:smth}

\subsection{Molecular Emission}
\label{sec:mole}

\subsection{Zero Level}
\label{sec:zp}

\subsubsection{Absolute Zero Level}
\label{sec:zp_abs}

\begin{center}
\begin{deluxetable*}{ccccccccc}
\tabletypesize{\scriptsize}
\tablecolumns{8} 
\tablewidth{0pc} 
\tablecaption{\label{table:offs} Input Map Properties \& Pre-processing (2015 release)} 
\tablehead{
\colhead{$\nu$ (GHz)} &
\colhead{Instrument(s)} &
\colhead{Offset ($K_{CMB}$)} & 
\colhead{Dipole ($K_{CMB}$)} & 
\colhead{$s_{857,\nu}$$\times$$u_{\nu}$} &
\colhead{$\sigma_{s_{857,\nu}}$$\times$$u_{\nu}$} &
\colhead{$n_{\nu}$} ($K_{CMB}$) &
\colhead{$c_{\nu}$} &
\colhead{FWHM ($'$)}
}
\startdata
100  & \PLANCK~HFI & 9.65$\times$10$^{-6}$$\pm$4.00$\times$10$^{-7}$ & 1.77$\times$10$^{-6}$    & 1.63$\times$10$^{-3}$ & 3.25$\times$10$^{-5}$  & 5.06$\times$10$^{-5}$ & 0.0009 & 9.68   \\
143  & \PLANCK~HFI & 2.34$\times$10$^{-5}$$\pm$7.72$\times$10$^{-7}$ & 1.48$\times$10$^{-6}$    & 4.79$\times$10$^{-3}$ & 9.58$\times$10$^{-5}$  & 2.18$\times$10$^{-5}$ & 0.0007 & 7.30   \\
217  & \PLANCK~HFI & 6.68$\times$10$^{-5}$$\pm$2.70$\times$10$^{-6}$ & 9.05$\times$10$^{-6}$    & 2.17$\times$10$^{-2}$ & 4.35$\times$10$^{-4}$  & 3.06$\times$10$^{-5}$ & 0.0016 & 5.02   \\
353  & \PLANCK~HFI & 4.01$\times$10$^{-4}$$\pm$2.06$\times$10$^{-5}$ & 3.09$\times$10$^{-5}$    & 9.90$\times$10$^{-2}$ & 1.98$\times$10$^{-3}$  & 1.01$\times$10$^{-4}$ & 0.0078 & 4.94   \\
     &             & Offset (MJy/sr)                                 & Dipole (MJy/sr)          & $s_{857,\nu}$         & $\sigma_{s_{857,\nu}}$ & $n_{\nu}$ (MJy/sr)    &        &        \\ \cline{3-7} \\ [-2ex]
545  & \PLANCK~HFI & 3.07$\times$10$^{-1}$$\pm$2.01$\times$10$^{-2}$ & 1.00$\times$10$^{-2}$    & 3.36$\times$10$^{-1}$ & 6.71$\times$10$^{-3}$  & 0.031                 & 0.061  & 4.83   \\
857  & \PLANCK~HFI & 4.72$\times$10$^{-1}$$\pm$6.00$\times$10$^{-2}$ & -                        & 1.0                   & 2.0$\times$10$^{-2}$   & 0.028                 & 0.064  & 4.64   \\
1250 & DIRBE       & 7.04$\times$10$^{-2}$$\pm$1.19$\times$10$^{-1}$ & -                        & 2.05                  & 4.10$\times$10$^{-2}$  & 0.42                  & 0.10   & 42     \\
2141 & DIRBE       & 9.94$\times$10$^{-2}$$\pm$1.53$\times$10$^{-1}$ & -                        & 2.65                  & 5.29$\times$10$^{-2}$  & 0.79                  & 0.10   & 42     \\
3000 & DIRBE/\IRAS & 0.0$\pm$4.3$\times$10$^{-2}$                    & -                        & 1.30                  & 2.60$\times$10$^{-2}$  & 0.06                  & 0.10   & \ 6.1
\enddata
\end{deluxetable*}
\end{center}

\subsubsection{Relative Zero Level}
\label{sec:relzero}

\section{Dust Emission Model}
\label{sec:modeling}

\section{Predicting the Observed SED}
\label{sec:bpcorr}

\section{Global Model Parameters}
\label{sec:global}

\section{MCMC Fitting Procedure}
\label{sec:fitting}

\subsection{Pixelization}
\label{sec:pix}

\subsection{Sampling Parameters}
\label{sec:samp}

\subsection{Markov Chains}
\label{sec:mcmc}

\subsection{Low-resolution Fits}
\label{sec:lores}

\subsection{Global Parameters Revisited}
\label{sec:hier}

\section{Optical Reddening}
\label{sec:ebv}

\subsection{Reddening Calibration Procedure}
\label{sec:calib_ebv}

\subsection{Rectifying the Reddening Residuals}

\section{Comparison of Emission Predictions}
\label{sec:em_compare}

\subsection{The 353-3000 GHz Frequency Range}
\label{sec:hifreq}

\subsection{The 100-217 GHz Frequency Range}
\label{sec:lofreq}

\section{Data Release}
\label{sec:release}

\section{Conclusions}
\label{sec:conclusion}

\subsection{Single-MBB versus Two-component emission}

\subsection{Towards a Replacement for SFD}
\label{sec:replace}

We gratefully acknowledge support from the National Science Foundation Graduate
Research Fellowship under Grant No. DGE1144152, and NASA grant NNX12AE08G. 
Based on observations obtained with Planck (http://www.esa.int/Planck), an ESA 
science mission with instruments and contributions directly funded by ESA 
Member States, NASA, and Canada. This research made use of the NASA 
Astrophysics Data System (ADS) and the IDL Astronomy User's Library at Goddard.
\footnote{Available at \texttt{http://idlastro.gsfc.nasa.gov}}

\begin{center}
\begin{deluxetable*}{llrrrrrrrrrr} 
\tabletypesize{\scriptsize}
\tablecolumns{11} 
\tablewidth{0pc} 
\tablecaption{\label{tab:global} Global Model Parameters (2015 release)} 
\tablehead{
\colhead{Number} &
\colhead{Model} &
\colhead{$f_1$} &
\colhead{$q_1/q_2$} & 
\colhead{$\beta_1$} & 
\colhead{$\beta_2$} &
\colhead{$T_2$} &
\colhead{$T_1$} &
\colhead{$n$} &
\colhead{D.O.F.} &
\colhead{$\chi^2$} &
\colhead{$\chi^2_{\nu}$}
}
\startdata
 1 & FDS99 best-fit  & 0.0363 & 13.0 & 1.67 & 2.70 & 15.67 & 9.12 & 1.055 & 7 & 55.0 & 7.86 \\
 2 & FDS99 general   & 0.0457 & 8.03 & 1.59 & 2.83 & 15.68 & 9.71 & 0.998 & 3 & 6.42 & 2.14 \\
 3 & single MBB      & 0.0    &  ... &  ... & 1.56 & 19.86 & ...  & 1.016 & 6 & 48.0 & 7.99 \\ %[-2ex]
\enddata
\end{deluxetable*}
\end{center}

\bibliographystyle{apj}
\bibliography{twocomp}

\end{document}