### 2015-06-22 ms.tex
filling in the introduction

\documentclass{emulateapj}

\usepackage[dvipdf]{epsfig} 
\usepackage[dvips]{rotating}
\usepackage{subfigure}
\usepackage{mathrsfs}
\usepackage{amsmath}

\newcommand{\IRAS}{{\it IRAS}}
\newcommand{\HERSCHEL}{{\it Herschel}}
\newcommand{\SPITZER}{{\it Spitzer}}
\newcommand{\PLANCK}{{\it Planck}}
\newcommand{\AKARI}{{\it AKARI}}
\newcommand{\COBE}{{\it COBE}}
\newcommand{\WISE}{{\it WISE}}

\bibpunct{(}{)}{;}{a}{}{,}

\shorttitle{\PLANCK~2015 Thermal Dust}

\shortauthors{Meisner \& Finkbeiner}

\begin{document}

\title{Modeling thermal dust emission with the {\it PLANCK} 2015 release maps}
\author{Aaron M. Meisner\altaffilmark{1,2}}
\author{Douglas P. Finkbeiner\altaffilmark{1,2}}
\altaffiltext{1}{Department of Physics, Harvard University, 17 Oxford Street, 
Cambridge, MA 02138, USA; ameisner@fas.harvard.edu}
\altaffiltext{2}{Harvard-Smithsonian Center for Astrophysics, 60 Garden St, 
Cambridge, MA 02138, USA; dfinkbeiner@cfa.harvard.edu}

\begin{abstract}

Based on the \PLANCK~2015 data release maps \citep{planck2015}, we present a 
thermal dust emission analysis analogous to that of \cite{meisner15}. We 
isolate thermal dust emission in the \PLANCK~2015 frequency maps, and present 
full-sky models of the unpolarized thermal dust emission using several 
paremeterizations of the dust spectrum, including single modified blackbody and
two-component models. The qualitative conclusions of \cite{meisner15} remain 
unchanged. In particular, zodiacal light still represents the dominant 
systematic problem on large angular scales, and the cosmic infrared background 
anisotropy remains the dominant systematic problem on small angular scales.

% mention frequency range 100-3000 GHz ?
% cite the FDS99 model in the intro as well ?
% mention that this isn't just a Planck-based SED, but includes DIRBE/IRAS ?
% mention that this is first ever thermal dust model based on Planck 2015 ?

\end{abstract}

\keywords{infrared: ISM, submillimeter: ISM, dust, extinction}

\section{Introduction}
% brief importance of dust
Interstellar dust in our Galaxy represents a crucial and pervasive foreground
for astronomical observations over an extremely wide range of photon energies.
In the optical, ultraviolet and near-infrared, extinction by intervening dust
attenuates and reddens the observed signal from background stars, galaxies and
transients. At millimeter wavelengths, thermal dust emission contaminates
observations of the cosmic microwave background (CMB).

% iras first revolution in mapping ISM cirrus
While the dust generally lies within a disk a few hundred parsecs in scale 
height centered about the Galactic midplane, its detailed distribution is very 
patchy and cloudlike, as first revealed in far-infrared thermal emission by the
groundbreaking \IRAS~mission \citep{low84, wheelock94}. Thus, for the sake of 
studying cosmological foregrounds and the Galactic interstellar medium (ISM), 
it is of great importance to map the distribution of dust over the entire sky, 
with the best possible angular resolution and sensitivity.

% planck more recent breakthrough in mapping dust emission due to high
%     angular resolution / sensitivity of broad range of frequencies
The \PLANCK~mission \citep{planck2013,planck2015} has recently provided a 
breakthrough in full-sky mapping of emission from interstellar dust, by 
obtaining relatively artifact free, full-sky maps in nine bandpasses from 30 
GHz to 857 GHz, with a superb combination of very high sensitivity and 
near-\IRAS~resolution from 217 GHz to 857 GHz. Although these \PLANCK~maps 
contain a superposition of various emission components, including the CMB, 
compact sources, and molecular emission, it is nevertheless possible to isolate
exquisite maps of thermal emission from interstellar dust grains.

% mention the existing dust data products, say they're all very useful for
%     many purposes
Several full-sky data products currently exist for predicting emission and/or
extinction by interstellar dust. For many years, the 
\citet[hereafter SFD]{SFD} dust map has been the industry standard for 
predicting Galactic reddening. The SFD reddening predictions rely on the
\IRAS~100$\mu$m dust emission map as a $6.1'$ FWHM template for dust optical 
depth, after modulating the emission by a $1.3^{\circ}$ FWHM temperature 
correction based on the DIRBE 140$\mu$m/240$\mu$m ratio, assuming the dust
emits as a $\beta$=2 modified blackbody (MBB). \citet[henceforth FDS99]{FDS99} 
combined the \IRAS/DIRBE data sets used by SFD with the FIRAS \citep{firas} 
dust spectra from 100-3000 GHz to extrapolate thermal dust emission from 
\IRAS~wavelengths to CMB frequencies. FDS99 found that no single MBB could
adequately fit the relatively sharp peak of the dust spectral energy 
distribution (SED) while also capturing the spectrum's flattening toward
millimeter wavelengths. FDS99 therefore fit the dust SED with a linear 
combination of two MBBs, which can be thought of as representing two 
distinct grain species in equilibrium with the same ISRF.

Both SFD/FDS99 DIRBE temperature limitation.
Planck allows us to get around temperature limitation. One sentence overview of
Planck team's model. One or two sentences about Meisner+Finkbeiner modeling.

% but say that they all still are not perfect, and we want to investigate
%     what effect swapping in the latest planck release has on the dust model
% and we also just want to have a model with data from latest Planck release,
%     which has never been done by anyone before
But none of these are yet perfect. Mention CIBA and zodi as examples from
Meisner + Finkbeiner. Also, no Planck 2015 release thermal dust model yet
exists at all.

% list of the sections
Much of our detailed methodology is identical to that of \cite{meisner15}. As a
result, throughout this paper, we mainly emphasize any notable differences 
relative to this previously published work, and otherwise refer the reader to 
\cite{meisner15} for thorough details.

In $\S$\ref{sec:data} we introduce the data used throughout this study. In
$\S$\ref{sec:prepro} we describe our preprocessing of the \PLANCK~maps to
isolate thermal emission from Galactic dust. In $\S$\ref{sec:modeling} we
explain the two-component emission model we apply to the \PLANCK-based dust
SED. In $\S$\ref{sec:bpcorr}, we discuss the details of predicting
\PLANCK~observations based on this dust model. In $\S$\ref{sec:global} we
derive constraints on our model's global parameters in light of the \PLANCK~HFI
maps. In $\S$\ref{sec:fitting} we detail the Markov chain Monte Carlo (MCMC)
method with which we have estimated the spatially varying parameters of our
model. In $\S$\ref{sec:ebv} we calibrate our derived optical depth to reddening
at optical wavelengths. In $\S$\ref{sec:em_compare} we compare our
two-component thermal dust emission predictions to those of \cite{planckdust}.
In $\S$\ref{sec:release} we present the full-sky maps of dust temperature and
optical depth we have obtained, and conclude in $\S$\ref{sec:conclusion}.

\section{Data}
\label{sec:data}

\section{Preprocessing}
\label{sec:prepro}

\subsection{CMB Anisotropy Removal}
\label{sec:cmb}

\subsection{Compact Sources}
\label{sec:ptsrc}

\subsection{Smoothing}
\label{sec:smth}

\subsection{Molecular Emission}
\label{sec:mole}

\subsection{Zero Level}
\label{sec:zp}

\subsubsection{Absolute Zero Level}
\label{sec:zp_abs}

\begin{center}
\begin{deluxetable*}{ccccccccc}
\tabletypesize{\scriptsize}
\tablecolumns{8} 
\tablewidth{0pc} 
\tablecaption{\label{table:offs} Input Map Properties \& Pre-processing (2015 release)} 
\tablehead{
\colhead{$\nu$ (GHz)} &
\colhead{Instrument(s)} &
\colhead{Offset ($K_{CMB}$)} & 
\colhead{Dipole ($K_{CMB}$)} & 
\colhead{$s_{857,\nu}$$\times$$u_{\nu}$} &
\colhead{$\sigma_{s_{857,\nu}}$$\times$$u_{\nu}$} &
\colhead{$n_{\nu}$} ($K_{CMB}$) &
\colhead{$c_{\nu}$} &
\colhead{FWHM ($'$)}
}
\startdata
100  & \PLANCK~HFI & 9.65$\times$10$^{-6}$$\pm$4.00$\times$10$^{-7}$ & 1.77$\times$10$^{-6}$    & 1.63$\times$10$^{-3}$ & 3.25$\times$10$^{-5}$  & 5.06$\times$10$^{-5}$ & 0.0009 & 9.68   \\
143  & \PLANCK~HFI & 2.34$\times$10$^{-5}$$\pm$7.72$\times$10$^{-7}$ & 1.48$\times$10$^{-6}$    & 4.79$\times$10$^{-3}$ & 9.58$\times$10$^{-5}$  & 2.18$\times$10$^{-5}$ & 0.0007 & 7.30   \\
217  & \PLANCK~HFI & 6.68$\times$10$^{-5}$$\pm$2.70$\times$10$^{-6}$ & 9.05$\times$10$^{-6}$    & 2.17$\times$10$^{-2}$ & 4.35$\times$10$^{-4}$  & 3.06$\times$10$^{-5}$ & 0.0016 & 5.02   \\
353  & \PLANCK~HFI & 4.01$\times$10$^{-4}$$\pm$2.06$\times$10$^{-5}$ & 3.09$\times$10$^{-5}$    & 9.90$\times$10$^{-2}$ & 1.98$\times$10$^{-3}$  & 1.01$\times$10$^{-4}$ & 0.0078 & 4.94   \\
     &             & Offset (MJy/sr)                                 & Dipole (MJy/sr)          & $s_{857,\nu}$         & $\sigma_{s_{857,\nu}}$ & $n_{\nu}$ (MJy/sr)    &        &        \\ \cline{3-7} \\ [-2ex]
545  & \PLANCK~HFI & 3.07$\times$10$^{-1}$$\pm$2.01$\times$10$^{-2}$ & 1.00$\times$10$^{-2}$    & 3.36$\times$10$^{-1}$ & 6.71$\times$10$^{-3}$  & 0.031                 & 0.061  & 4.83   \\
857  & \PLANCK~HFI & 4.72$\times$10$^{-1}$$\pm$6.00$\times$10$^{-2}$ & -                        & 1.0                   & 2.0$\times$10$^{-2}$   & 0.028                 & 0.064  & 4.64   \\
1250 & DIRBE       & 7.04$\times$10$^{-2}$$\pm$1.19$\times$10$^{-1}$ & -                        & 2.05                  & 4.10$\times$10$^{-2}$  & 0.42                  & 0.10   & 42     \\
2141 & DIRBE       & 9.94$\times$10$^{-2}$$\pm$1.53$\times$10$^{-1}$ & -                        & 2.65                  & 5.29$\times$10$^{-2}$  & 0.79                  & 0.10   & 42     \\
3000 & DIRBE/\IRAS & 0.0$\pm$4.3$\times$10$^{-2}$                    & -                        & 1.30                  & 2.60$\times$10$^{-2}$  & 0.06                  & 0.10   & \ 6.1
\enddata
\end{deluxetable*}
\end{center}

\subsubsection{Relative Zero Level}
\label{sec:relzero}

\section{Dust Emission Model}
\label{sec:modeling}

\begin{center}
\begin{deluxetable*}{llrrrrrrrrrr} 
\tabletypesize{\scriptsize}
\tablecolumns{11} 
\tablewidth{0pc} 
\tablecaption{\label{tab:global} Global Model Parameters (2015 release)} 
\tablehead{
\colhead{Number} &
\colhead{Model} &
\colhead{$f_1$} &
\colhead{$q_1/q_2$} & 
\colhead{$\beta_1$} & 
\colhead{$\beta_2$} &
\colhead{$T_2$} &
\colhead{$T_1$} &
\colhead{$n$} &
\colhead{D.O.F.} &
\colhead{$\chi^2$} &
\colhead{$\chi^2_{\nu}$}
}
\startdata
 1 & FDS99 best-fit  & 0.0363 & 13.0 & 1.67 & 2.70 & 15.67 & 9.12 & 1.055 & 7 & 55.0 & 7.86 \\
 2 & FDS99 general   & 0.0457 & 8.03 & 1.59 & 2.83 & 15.68 & 9.71 & 0.998 & 3 & 6.42 & 2.14 \\
 3 & single MBB      & 0.0    &  ... &  ... & 1.56 & 19.86 & ...  & 1.016 & 6 & 48.0 & 7.99 \\ %[-2ex]
\enddata
\end{deluxetable*}
\end{center}

\section{Predicting the Observed SED}
\label{sec:bpcorr}

\section{Global Model Parameters}
\label{sec:global}

\section{MCMC Fitting Procedure}
\label{sec:fitting}

\subsection{Pixelization}
\label{sec:pix}

\subsection{Sampling Parameters}
\label{sec:samp}

\subsection{Markov Chains}
\label{sec:mcmc}

\subsection{Low-resolution Fits}
\label{sec:lores}

\subsection{Global Parameters Revisited}
\label{sec:hier}

\section{Optical Reddening}
\label{sec:ebv}

\subsection{Reddening Calibration Procedure}
\label{sec:calib_ebv}

\subsection{Rectifying the Reddening Residuals}

\section{Comparison of Emission Predictions}
\label{sec:em_compare}

\subsection{The 353-3000 GHz Frequency Range}
\label{sec:hifreq}

\subsection{The 100-217 GHz Frequency Range}
\label{sec:lofreq}

\section{Data Release}
\label{sec:release}

\section{Conclusions}
\label{sec:conclusion}

\subsection{Single-MBB versus Two-component emission}

\subsection{Towards a Replacement for SFD}
\label{sec:replace}

We gratefully acknowledge support from the National Science Foundation Graduate
Research Fellowship under Grant No. DGE1144152, and NASA grant NNX12AE08G. 
Based on observations obtained with Planck (http://www.esa.int/Planck), an ESA 
science mission with instruments and contributions directly funded by ESA 
Member States, NASA, and Canada. This research made use of the NASA 
Astrophysics Data System (ADS) and the IDL Astronomy User's Library at Goddard.
\footnote{Available at \texttt{http://idlastro.gsfc.nasa.gov}}

\bibliographystyle{apj}
\bibliography{twocomp}

\end{document}